---
title: "Moeller data 2024 Ongoing"
author: "Ian Bach"
date: "2024-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The dataset utilized in the "Moeller data" project comprises comprehensive baseball statistics from Moeller, focusing on a range of variables significant to understanding game dynamics and player performance. 

It includes detailed entries for each pitch, annotated with information about the batter, the count (balls and strikes), the pitch result (e.g., "Strike Looking"), and which team the batter belongs to. 

The data is stored in a CSV format, ensuring easy accessibility and manipulation using R programming tools. This dataset is particularly tailored for analyzing specific in-game scenarios, such as the incidence of strikes when the count is favorable to the batter. 

By examining these detailed metrics, the project aims to derive insights that could influence team strategies and player coaching, offering a granular view of performance during crucial game moments.

```{r Moeller Data}
# Load required libraries
library(dplyr)   # For data manipulation
library(ggplot2) # For data visualization
library(readr)   # Enhanced data reading capabilities
library(scales)  # For percentage formatting
library(gridExtra)  # For arranging multiple plots
library(purrr)
library(tidyr)
# Set the working directory
setwd("C:/Users/ibach/OneDrive - Terillium/Pictures/Moller Misc/Updated_data")

# Read the CSV file
df <- read.csv('UPDATED_Moeller_Baseball_2024.csv')

# View the data frame
head(df)
```

```{r Analysis of Strike Looking Incidence for Moeller Batters with Specific Count and Strike Scenario, echo=FALSE}
# Filter the original dataframe to include only rows where the batter has 1 ball, 0 strikes, and is from the 'Moeller' team
filtered_df <- df %>%
  filter(Balls == 1, Strikes == 0, BatterTeam == 'Moeller')

# Count the number of pitches that resulted in a 'Strike Looking' from the filtered dataframe
strike_looking_count <- sum(filtered_df$PitchResult == 'Strike Looking', na.rm = TRUE)

# Count the total number of entries in the filtered dataframe
total_count <- nrow(filtered_df)

# Create percentage looking equation
percentage_strike_looking <- (strike_looking_count / total_count) * 100

# Output the shape of the filtered dataframe (number of rows, number of columns)
dim(filtered_df)

```
```{r Calculation of Specific Pitch and At-Bat Outcomes in Filtered Data}
# Assuming dplyr is already loaded and filtered_df has been defined as per your previous request.

# Create Strike Swinging count by filtering Pitch Results = 'Strike Swing and Miss'
strike_swinging_count <- filtered_df %>%
  filter(PitchResult == "Strike Swing and Miss") %>%
  nrow()

# Filtered count for pitch result to 'Strike Foul'
strike_foul_count <- filtered_df %>%
  filter(PitchResult == "Strike Foul") %>%
  nrow()

# Filtered count for pitch result to 'Strike In Play'
strike_inplay_count <- filtered_df %>%
  filter(PitchResult == "Strike In Play") %>%
  nrow()

# Filtered count for pitch result as 'Ball'
ball_count <- filtered_df %>%
  filter(PitchResult == "Ball") %>%
  nrow()

# Filtered count for AtBatResult as 1B, 2B, 3B, HR
hit_count <- filtered_df %>%
  filter(AtBatResult %in% c('1B', '2B', '3B', 'HR')) %>%
  nrow()

```


```{r Calculation of Pitch Outcome Percentages from Total Pitch Counts}
# Calculate the percentage of Strike Swinging
percentage_strike_swinging <- (strike_swinging_count / total_count) * 100

# Calculate the percentage of Strike Foul
percentage_strike_foul <- (strike_foul_count / total_count) * 100

# Calculate the percentage of Strike In Play
percentage_strike_inplay <- (strike_inplay_count / total_count) * 100

# Calculate the percentage of Balls
percentage_ball <- (ball_count / total_count) * 100

# Calculate the percentage of Hits
percentage_hit <- (hit_count / total_count) * 100

```


```{r Visualization of Pitch Outcome Percentages in a 1-0 Count for the 2024 Season}
# Load the ggplot2 library for visualization
library(ggplot2)

# Create a data frame for plotting
data <- data.frame(
  Category = c('Strike Swinging', 'Strike Foul', 'Strike In Play', 'Strike Looking', 'Ball', 'Hit'),
  Count = c(percentage_strike_swinging, percentage_strike_foul, percentage_strike_inplay, percentage_strike_looking, percentage_ball, percentage_hit),
  Fill = rep(c('navy', 'gold'), each = 3)
)

# Create the bar graph
ggplot(data, aes(x = Category, y = Count, fill = Fill)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_identity() +
  labs(title = "1-0 Count Outcome % 2024 Season", x = "Outcome", y = "%") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Adjust x labels for better readability

```
```{r Creation of DF and Visualization to create Pitch Result % for Moeller Pitchers}
# Extract the unique values from the 'PitchResult' column of the dataframe to see all possible outcomes
unique_pitch_results <- unique(df$PitchResult)

# Creation of Strike events
strike_events <- c('Strike Looking', 'Strike Swing and Miss', 'Strike Foul')

# Add a new column 'is_strike' initialized to 0 and set to 1 where 'PitchResult' is in the list of strike events
df <- df %>%
  mutate(is_strike = ifelse(PitchResult %in% strike_events, 1, 0))

# Initialize a new column 'is_moeller_pitcher' with 0 and set to 1 where 'PitcherTeam' is 'Moeller'
df <- df %>%
  mutate(is_moeller_pitcher = ifelse(PitcherTeam == 'Moeller', 1, 0))

# Create a dataframe to calculate the mean of 'is_strike' for each 'Pitcher' and 'PitchType' for Moeller pitchers
df1 <- df %>%
  filter(is_moeller_pitcher == 1) %>%
  group_by(Pitcher, PitchType) %>%
  summarise(Strike_Percentage = mean(is_strike, na.rm = TRUE) * 100) %>%
  ungroup() %>%
  mutate(Strike_Percentage = round(Strike_Percentage, 2))

# Create another dataframe to count occurrences of each 'PitchType' for Moeller pitchers
df2 <- df %>%
  filter(is_moeller_pitcher == 1) %>%
  group_by(Pitcher, PitchType) %>%
  summarise(Count = n()) %>%
  ungroup()

# Merge the two dataframes on 'Pitcher' and 'PitchType'
df3 <- merge(df1, df2, by = c("Pitcher", "PitchType"))

# Print the merged dataframe
print(df3)

```
```{r Generation of Individual Bar Plots for Strike Percentages by Pitch Type for Each Pitcher}
# Loop through each unique pitcher in 'df3'
unique_pitchers <- unique(df3$Pitcher)

for (pitcher in unique_pitchers) {
  print(pitcher)  # Print the pitcher's name
  
  # Create a data frame 'df_plt' containing data for the current pitcher in the loop
  df_plt <- filter(df3, Pitcher == pitcher)
  
  # Create barplot
  p <- ggplot(df_plt, aes(x = Strike_Percentage, y = reorder(PitchType, Strike_Percentage), fill = PitchType)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste('2024 Season Strike Percentage by Pitch Type -', pitcher), x = "Strike Percentage", y = "Pitch Type") +
    scale_x_continuous(labels = percent_format(scale = 1)) +  # Assuming the values are already in percentages
    theme_minimal()
  
  print(p)  # Display the plot
}

```

```{r Faceted Bar Chart Visualization of Strike Percentages by Pitch Type for Each Pitcher}


```

```{r Creation of a new DF called Offspeed:}
# Organize pitch type data to include only 'Change Up' and 'Breaking Ball' for Moeller Hitters
offspeed <- df %>%
  filter(PitchType %in% c('Change Up', 'Breaking Ball'), BatterTeam == 'Moeller')

# Print the top entries of the offspeed data frame
print(head(offspeed))

```

```{r Counting Specific Pitch Outcomes for Offspeed Pitches}
# Count on Strike Swing and Miss
offstrike_swinging_count <- offspeed %>%
  filter(PitchResult == 'Strike Swing and Miss') %>%
  nrow()

# Count on Strike Looking
offstrike_looking_count <- offspeed %>%
  filter(PitchResult == 'Strike Looking') %>%
  nrow()

# Count on Strike Foul
offstrike_foul_count <- offspeed %>%
  filter(PitchResult == 'Strike Foul') %>%
  nrow()

# Count on Strike In Play
offstrike_inplay_count <- offspeed %>%
  filter(PitchResult == 'Strike In Play') %>%
  nrow()

# Count on Ball
offball_count <- offspeed %>%
  filter(PitchResult == 'Ball') %>%
  nrow()

# Count on results 1B, 2B, 3B, and HR
offhit_count <- offspeed %>%
  filter(AtBatResult %in% c('1B', '2B', '3B', 'HR')) %>%
  nrow()

# Total count of offspeed pitches
offtotal_count <- nrow(offspeed)

```


```{r Calculation of Pitch Outcome Percentages for Offspeed Pitches}
if(offtotal_count == 0) {
  print("Total count cannot be zero.")
} else {
 # Calculate percentages
offpercentage_strike_swinging <- (offstrike_swinging_count / offtotal_count) * 100
offpercentage_strike_looking <- (offstrike_looking_count / offtotal_count) * 100
offpercentage_strike_foul <- (offstrike_foul_count / offtotal_count) * 100
offpercentage_strike_inplay <- (offstrike_inplay_count / offtotal_count) * 100
offpercentage_ball <- (offball_count / offtotal_count) * 100
offpercentage_hit <- (offhit_count / offtotal_count) * 100

}

# Print the results
print(paste("Strike Swinging Percentage:", offpercentage_strike_swinging))
print(paste("Strike Looking Percentage:", offpercentage_strike_looking))
print(paste("Strike Foul Percentage:", offpercentage_strike_foul))
print(paste("Strike In Play Percentage:", offpercentage_strike_inplay))
print(paste("Offspeed Ball Percentage:", offpercentage_ball))
print(paste("Offspeed Hit Percentage:", offpercentage_hit))

```

```{r Visualization of Pitch Outcomes for Moeller Hitters in the 2024 Season}
# Define the categories and counts
categories <- c('Strike Swinging', 'Strike Foul', 'Strike In Play', 'Strike Looking', 'Ball', 'Hit')
counts <- c(offpercentage_strike_swinging, offpercentage_strike_foul, offpercentage_strike_inplay, offpercentage_strike_looking, offpercentage_ball, offpercentage_hit)

# Create a data frame for plotting
data <- data.frame(Category = categories, Count = counts)

# Create a bar chart
ggplot(data, aes(x = Category, y = Count, fill = Category)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = sprintf("%.2f%%", Count)), vjust = -0.3, size = 3.5) +
  scale_fill_manual(values = c('navy', 'gold', 'navy', 'gold', 'navy', 'gold')) +
  labs(title = "Breaking Ball/Changeup Outcome (Moeller Hitters) % 2024 Season",
       x = "Outcome",
       y = "%") +
  theme_minimal()

```

```{r Analysis of Pitcher Attack Zone Effectiveness and Frequency for Moeller Pitchers}
# Creation of DF to Mean Attack Zone
df4 <- df %>%
  filter(is_moeller_pitcher == 1) %>%
  group_by(Pitcher, AttackZone) %>%
  summarise(Strike_Percent = mean(is_strike, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Strike_Percent = round(Strike_Percent * 100, 2))

# Creation of DF to Count Attack Zone
df5 <- df %>%
  filter(is_moeller_pitcher == 1) %>%
  group_by(Pitcher, AttackZone) %>%
  summarise(Count = n()) %>%
  ungroup()

# Merge DF4 with DF5 to create df_attack
df_attack <- merge(df4, df5, by = c("Pitcher", "AttackZone"))

# Removing rows with blank AttackZone
df_attack <- df_attack %>%
  filter(AttackZone != "" & !is.na(AttackZone))

# Print the final data frame
print(df_attack)


```
```{r Faceted Visualization of Strike Percentages by Attack Zone for Each Pitcher}
# Ensure the data is correct
df_attack$Pitcher <- as.factor(df_attack$Pitcher)  # Ensure 'Pitcher' is a factor

# Create the plot
g <- ggplot(df_attack, aes(x = Strike_Percent, y = reorder(AttackZone, Strike_Percent))) +
  geom_bar(stat = "identity", fill = "blue") +
  facet_wrap(~Pitcher, scales = "free", ncol = 4) +
  scale_x_continuous(labels = percent_format(scale = 1), 
                     breaks = seq(0, 100, 10)) +  # Set breaks at every 10%
  labs(title = "2024 Season Strike Percentage by Attack Zone",
       x = "Strike Percentage",
       y = "Attack Zone") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  # Adjust text for better visibility

# Print the plot
print(g)


```

```{r Analysis of Pitch Results for Moeller Team Batters Against Left-Handed Pitchers Non Facet}

# Ensure the data is correct
df_attack$Pitcher <- as.factor(df_attack$Pitcher)

# Get unique pitchers
pitchers <- unique(df_attack$Pitcher)

# Create a list to store ggplot objects
plot_list <- list()

# Iterate over each pitcher to create plots
for (pitcher in pitchers) {
  df_plt <- filter(df_attack, Pitcher == pitcher)  # Filter data for the pitcher
  p <- ggplot(df_plt, aes(x = Strike_Percent, y = reorder(AttackZone, Strike_Percent))) +
    geom_bar(stat = "identity", fill = "blue") +
    labs(title = paste("2024 Season Strike Percentage by Attack Zone -", pitcher),
         x = "Strike Percentage",
         y = "Attack Zone") +
    theme_minimal() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
  print(p)
}


```

```{r Comprehensive Analysis of Pitch Results for Moeller Batters Against Left-Handed Pitchers}

# Initialize a new column 'is_moeller_hitter' in the dataframe 'df' and set its default value to 0
df$is_moeller_hitter <- 0

# Update this column to 1 for rows where the 'BatterTeam' is 'Moeller'
df$is_moeller_hitter[df$BatterTeam == "Moeller"] <- 1

# Retrieve and display the unique values from the 'PitchResult' column
unique(df$PitchResult)

# Filter 'df' to select various conditions and create summaries
# Use a function to streamline the process
create_summary <- function(result) {
  df %>%
    filter(is_moeller_hitter == 1, PitcherHand == "L", PitchResult == result) %>%
    count(Batter, name = result)
}

# Creating individual summaries for each type of pitch result
df_strike_looking <- create_summary("Strike Looking")
df_strike_in_play <- create_summary("Strike In Play")
df_ball <- create_summary("Ball")
df_whiff <- create_summary("Strike Swing and Miss")
df_foul <- create_summary("Strike Foul")

# Merge multiple dataframes sequentially using 'Batter' as the key column
df_merged <- reduce(list(df_strike_looking, df_strike_in_play, df_ball, df_whiff, df_foul), 
                    full_join, by = "Batter")

# Print the final merged dataframe
print(df_merged)

```

```{r Hitter Visualization Comprehensive Analysis of Pitch Results for Moeller Batters Against Left-Handed Pitchers}

# Assuming 'df_merged' is your final dataframe
# Check if the data needs to be reshaped from wide to long format
df_long <- df_merged %>%
  pivot_longer(cols = c('Strike Looking', 'Strike In Play', 'Ball', 'Strike Swing and Miss', 'Strike Foul'),
               names_to = "PitchResult", values_to = "Count")

# Create a list to store ggplot objects, if you want to plot them all at once later
plot_list <- list()

# Create a plot for each batter
batters <- unique(df_long$Batter)
for (batter in batters) {
  # Filter the data for one batter
  df_batter <- filter(df_long, Batter == batter)

  # Generate the plot
  p <- ggplot(df_batter, aes(x = PitchResult, y = Count, fill = PitchResult)) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    labs(title = paste("Pitch Result Counts for", batter),
         x = "Pitch Result Type",
         y = "Count") +
    scale_fill_brewer(palette = "Set3")  # Color palette

  # Print the plot directly
  print(p)
  
  # Store the plot in the list
  plot_list[[batter]] <- p
}

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
